---
title: "Experiment 2 Analyses"
author: "Kushin Mukherjee"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(psych)
library(stringr)
library(proxy)

# Load the package
library(osfr)
```

```{r}

analysis_dir<- getwd()

data_dir <-paste0(analysis_dir,'/../data/exp_2')
### if data_dir doesn't exist create it
if (!dir.exists(data_dir)) {
  dir.create(data_dir)
}

results_dir <-paste0(analysis_dir,'/../results/')
### if results_dir doesn't exist create it
if (!dir.exists(results_dir)) {
  dir.create(results_dir)
}

plot_dir <- paste0(analysis_dir,'/../plots/')

### if plot_dir doesn't exist create it
if (!dir.exists(plot_dir)) {
  dir.create(plot_dir)
}

```

### OSF Download if needed

```{r}

osf_token <- readLines("../km_osf_pat.txt", warn = FALSE)[1]
# Authenticate with OSF (you'll be prompted to generate a PAT token)
osf_auth(token=osf_token)

# Retrieve the OSF component
osf_project <- osf_retrieve_node("6yqhd")

# List all files in the component
files <- osf_ls_files(osf_project, n_max = Inf)


# Download all files
osf_download(files, path = data_dir, conflicts = "overwrite")


```

### Load data files

```{r}
data_files <-list.files(path = data_dir, full.names = TRUE, pattern="*.csv")
print(paste0('we have data from ',length(data_files),' participants'))



df_list <- lapply(data_files, function(f) {
  df <- read.csv(f)
  df[, !names(df) %in% "SONA_id", drop = FALSE]
})

experiment_data_raw<- bind_rows(df_list)

ratings_trials<- experiment_data_raw %>% filter(trial_type== 'html-slider-response')

ratings_trials_exp<- ratings_trials %>% filter(!practiceTrial %in% c('True','true'))

### filter ratings_trials_exp such that we have only 1 subject_id per condition_num
# set.seed(56)  
# ratings_trials_exp <- ratings_trials_exp %>%
#   group_by(condition_num) %>%
#   filter(subject_id %in% sample(unique(subject_id), 1)) %>%
#   ungroup()


ratings_trials_exp$response<- as.numeric(ratings_trials_exp$response)
ratings_trials_exp$response<- (ratings_trials_exp$response + 200)/400


ratings_trials_exp_grouped <- ratings_trials_exp%>%group_by(concept,scale_name,mapping_code, shift_code)%>%summarize(mean_rating = mean(response))


### get the number of unique participants per condition in ratings_trials_exp
subject_dist<-ratings_trials_exp %>% 
  group_by(condition_num) %>% 
  summarize(num_participants = n_distinct(subject_id))

subject_dist

### can you list all the numbers between 0 and 79 not present in subject_dist$condition_num
missing_conditions <- setdiff(0:40, subject_dist$condition_num)
print(paste0('missing conditions: ', paste(missing_conditions, collapse = ', ')))



write.csv(ratings_trials_exp_grouped, '../data/exp2_ratings_trials_exp_grouped.csv')

```

```{r}
library(ggimage)

preds_per_color <- read.csv(paste0(results_dir, "matplotlib_colorscales_w_predictions_2.csv"),
                             stringsAsFactors = FALSE)
# 
# analysis_df <- ratings_trials_exp_grouped%>%inner_join(predicted_ratings_grouped_uniform_long, by = c("concept", "scale_name"))
# 
# analysis_df$path_name <- paste0("../images/small/map02_", analysis_df$scale_name, "_scaleDef_orientDef_000_25px.png")
# # 2. Build the plot
# p <- ggplot(analysis_df, aes(x = mean_prediction, y = mean_rating)) +
#   geom_image(aes(image = path_name), size = 0.05) +
#   geom_smooth(method = 'lm', se = FALSE, color = '#06c7bd') +
#   facet_wrap(~concept, nrow = 1) +
#   labs(x = 'Predicted Ratings', y = 'Human Ratings') +
#   theme_linedraw() +
#   scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
#   scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
#   coord_fixed(ratio = 1) +
#   theme(
#     panel.spacing.y = unit(1, "lines"),
#     panel.spacing.x = unit(2, "lines"),
#     plot.margin = margin(10, 20, 10, 10, "pt"),
#     axis.title = element_text(size = 12),
#     axis.text = element_text(size = 12),
#      strip.text = element_text(size = 12)
#   )
# 
# 
# # 3. Save the result
# n_facets <- length(unique(analysis_df$concept))
# panel_size <- 3.5  # inches per panel
# ggsave(
#   paste0(plot_dir, 'exp_2_scatterplot_predicted_vs_observed_ratings.pdf'),
#   plot = p,
#   width = panel_size * n_facets + 1,  # +1 for margins/labels
#   height = panel_size + 1.5,           # +1.5 for axis labels/title
#   dpi = 300
# )
# 
# p
```

### Uniform Predictions (equal-weight baseline, as in Exp1)

Assumes all 256 colors in a colorscale are present equally — the same approach used in Exp1. This serves as a baseline to compare against the proportion-weighted predictions below.

```{r}
# ── Uniform (equal-weight) predictions per colorscale ─────────────────────────

uniform_preds <- preds_per_color %>%
  group_by(colorscale) %>%
  summarize(
    angry   = mean(angry_pred),
    fearful = mean(fearful_pred),
    happy   = mean(happy_pred),
    sad     = mean(sad_pred),
    disgust = mean(disgust_pred),
    .groups = "drop"
  ) %>%
  rename(scale_name = colorscale) %>%
  pivot_longer(c(angry, fearful, happy, sad, disgust),
               names_to = "concept", values_to = "uniform_pred")

# ── Join with observed ratings and collapse to 40 points per emotion ──────────
# scaleDef+darkshift ≡ scaleRev+lightshift (both show high-end default colors)
# scaleDef+lightshift ≡ scaleRev+darkshift (both show low-end default colors)

analysis_uniform <- ratings_trials_exp_grouped %>%
  inner_join(uniform_preds, by = c("concept", "scale_name")) %>%
  mutate(condition = ifelse(
    (mapping_code == "scaleDef" & shift_code == "darkshift") |
    (mapping_code == "scaleRev" & shift_code == "lightshift"),
    "high_end", "low_end"
  )) %>%
  group_by(concept, scale_name, condition) %>%
  summarize(
    mean_rating  = mean(mean_rating),
    uniform_pred = mean(uniform_pred),
    .groups = "drop"
  ) %>%
  mutate(path_name = paste0(
    "../images/shifted_maps_small/map02_",
    scale_name, "_scaleDef_orientDef_000_",
    ifelse(condition == "high_end", "darkshift", "lightshift"), ".png"
  ))
```

```{r}
# ── Plot: uniform predicted vs observed, faceted concept × shift ──────────────

p_uniform <- ggplot(analysis_uniform, aes(x = uniform_pred, y = mean_rating)) +
  geom_image(aes(image = path_name), size = 0.06) +
  geom_smooth(method = "lm", se = FALSE, color = "#06c7bd") +
  facet_grid( cols = vars(concept)) +
  labs(x = "Uniform Predicted Rating", y = "Human Rating") +
  theme_linedraw() +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
  coord_fixed(ratio = 1) +
  theme(
    panel.spacing.x = unit(2, "lines"),
    panel.spacing.y = unit(1, "lines"),
    plot.margin     = margin(10, 20, 10, 10, "pt"),
    axis.title      = element_text(size = 12),
    axis.text       = element_text(size = 12),
    strip.text      = element_text(size = 12)
  )

n_concepts <- length(unique(analysis_uniform$concept))
panel_size <- 3.5

ggsave(
  paste0(plot_dir, "exp_2_scatterplot_uniform_pred_vs_observed.pdf"),
  plot   = p_uniform,
  width  = panel_size * n_concepts + 1,
  height = panel_size * 2 + 2,
  dpi    = 300
)

p_uniform
```

```{r}
# ── Correlation: uniform predictions vs observed (per concept) ────────────────

uniform_corrs_by_concept <- analysis_uniform %>%
  group_by(concept) %>%
  summarize(
    r    = cor(uniform_pred, mean_rating, use = "complete.obs"),
    n    = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(r))

uniform_corrs_by_concept
```

### Weighted Predictions (color-proportion-weighted)

Exp1 predictions assumed all 256 colors in a scale appear equally. Here we weight each color's predicted affect rating by the proportion of map pixels that actually display that color. Proportions were pre-computed per (map_id × scale_name × mapping_code) by `prep/compute_color_proportions.R` and are orientation/rotation/shift invariant (same pixels, different arrangement or brightness). The join to human data is at the full stimulus grain: map × scale × mapping × rotation × shift (3200 unique stimuli), with weighted_pred broadcast across rotation and shift variants.

```{r}
# ── Load proportions ──────────────────────────────────────────────────────────

proportions <- read.csv(paste0(results_dir, "color_proportions.csv"),
                        stringsAsFactors = FALSE)

# ── Pivot predictions to long format ──────────────────────────────────────────

concept_cols <- c("angry_pred", "happy_pred", "fearful_pred", "disgust_pred", "sad_pred")

preds_long <- preds_per_color %>%
  select(colorscale, index, all_of(concept_cols)) %>%
  pivot_longer(all_of(concept_cols), names_to = "concept", values_to = "pred") %>%
  mutate(concept = str_remove(concept, "_pred"))

# ── Compute weighted predictions ───────────────────────────────────────────────
# proportion × pred, summed over index, per stimulus × concept

weighted_preds <- proportions %>%
  left_join(preds_long, by = c("scale_name" = "colorscale", "index")) %>%
  group_by(map_id, scale_name, mapping_code, rotation_code, shift_code, concept) %>%
  summarize(weighted_pred = sum(proportion * pred), .groups = "drop")

weighted_preds
```

```{r}
# ── Fine-grained observed ratings ─────────────────────────────────────────────
# Group at the unique-stimulus level: map × scale × mapping × rotation × shift.
# Orientation is collapsed (mirroring preserves color proportions and counts).
# rotation_code in the data uses bare integers ("0","90",...); filenames use
# zero-padded strings ("000","090",...) — conversion handled in path construction.

ratings_fine <- ratings_trials_exp %>%
  mutate(map_id = as.integer(map_id)) %>%
  group_by(concept, map_id, scale_name, mapping_code, rotation_code, shift_code) %>%
  summarize(mean_rating = mean(response), .groups = "drop")

# ── Join observed ratings with weighted predictions ────────────────────────────

analysis_weighted <- ratings_fine %>%
  left_join(weighted_preds,
            by = c("map_id", "scale_name", "mapping_code",
                   "rotation_code", "shift_code", "concept"))

analysis_weighted
```

```{r}
# ── Plot: weighted predicted vs observed ratings, faceted by concept ───────────

# Collapse to 40 points per emotion (20 scales × 2 conditions)
# scaleDef+darkshift ≡ scaleRev+lightshift (both show high-end default colors)
# scaleDef+lightshift ≡ scaleRev+darkshift (both show low-end default colors)
analysis_grouped <- analysis_weighted %>%
  mutate(condition = ifelse(
    (mapping_code == "scaleDef" & shift_code == "darkshift") |
    (mapping_code == "scaleRev" & shift_code == "lightshift"),
    "high_end", "low_end"
  )) %>%
  group_by(concept, scale_name, condition) %>%
  summarize(
    mean_human_rating     = mean(mean_rating),
    mean_model_prediction = mean(weighted_pred),
    .groups = "drop"
  ) %>%
  mutate(path_name = paste0(
    "../images/shifted_maps_small/map02_",
    scale_name, "_scaleDef_orientDef_000_",
    ifelse(condition == "high_end", "darkshift", "lightshift"), ".png"
  ))

p_weighted <- ggplot(analysis_grouped, aes(x = mean_model_prediction, y = mean_human_rating)) +
  geom_image(aes(image = path_name), size = 0.06) +
  geom_smooth(method = "lm", se = FALSE, color = "#06c7bd") +
  facet_grid( cols = vars(concept)) +
  labs(x = "Weighted Predicted Rating", y = "Human Rating") +
  theme_linedraw() +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
  coord_fixed(ratio = 1) +
  theme(
    panel.spacing.x  = unit(2, "lines"),
    panel.spacing.y  = unit(1, "lines"),
    plot.margin      = margin(10, 20, 10, 10, "pt"),
    axis.title       = element_text(size = 12),
    axis.text        = element_text(size = 12),
    strip.text       = element_text(size = 12)
  )

n_concepts <- length(unique(analysis_grouped$concept))
panel_size <- 3.5

ggsave(
  paste0(plot_dir, "exp_2_scatterplot_weighted_pred_vs_observed.pdf"),
  plot   = p_weighted,
  width  = panel_size * n_concepts + 1,
  height = panel_size * 2 + 2,
  dpi    = 300
)

p_weighted
```

```{r}
# ── Correlation: weighted predictions vs observed (per concept) ────────────────

# analysis_weighted %>%
#   group_by(concept) %>%
#   summarize(
#     r     = cor(weighted_pred, mean_rating, use = "complete.obs"),
#     r_sq  = r^2,
#     n     = n(),
#     .groups = "drop"
#   ) %>%
#   arrange(desc(r))


weighted_corrs_by_concept <- analysis_grouped %>%
  group_by(concept) %>%
  summarize(
    r     = cor(mean_human_rating, mean_model_prediction, use = "complete.obs"),
    n     = n(),

    .groups = "drop"
  ) %>%
  arrange(desc(r))

weighted_corrs_by_concept
```

### Comparing weighted vs. uniform correlations (Fisher r-to-z)

Aggregate-level correlations (40 scale × condition points per concept) for the weighted and uniform models are compared per concept using a Fisher r-to-z test. Note this test assumes the two correlations are independent, which is not strictly true here since both are computed against the same underlying human ratings — treat these as a first pass; a dependent-correlations test (e.g. the Steiger-recommended Williams t-test / `cocor`) is the more appropriate follow-up.

One-tailed, testing the directional hypothesis that the weighted model correlates more strongly with human ratings than the uniform model (r1 is assumed to be the weighted correlation).

```{r}
# ── Fisher r-to-z test for two (treated-as-independent) correlations ─────────
# One-tailed: tests r1 > r2 (call with r1 = weighted, r2 = uniform)

fisher_r_to_z_test <- function(r1, n1, r2, n2) {
  z1 <- atanh(r1)
  z2 <- atanh(r2)
  se <- sqrt(1 / (n1 - 3) + 1 / (n2 - 3))
  z  <- (z1 - z2) / se
  p  <- 1 - pnorm(z)
  tibble(z_diff = z, p_value = p)
}

corr_comparison <- uniform_corrs_by_concept %>%
  rename(r_uniform = r, n_uniform = n) %>%
  inner_join(
    weighted_corrs_by_concept %>% rename(r_weighted = r, n_weighted = n),
    by = "concept"
  ) %>%
  rowwise() %>%
  mutate(fisher_r_to_z_test(r_weighted, n_weighted, r_uniform, n_uniform)) %>%
  ungroup() %>%
  mutate(
    sig = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01  ~ "**",
      p_value < 0.05  ~ "*",
      TRUE ~ "ns"
    )
  ) %>%
  arrange(p_value)

corr_comparison
```

### Comparing weighted vs. uniform correlations (Steiger-recommended Williams t-test)

The Fisher r-to-z test above treats `r_weighted` and `r_uniform` as independent, but they aren't: both are computed against the same 40 human-rating points, just correlated with two different predictors (weighted_pred and uniform_pred). This is the "overlapping/dependent correlations" case (two correlations sharing one variable), for which Steiger (1980) recommends the Williams (1959) modification of Hotelling's t-test over the naive Fisher r-to-z test. It additionally requires `r23`, the correlation *between* the two predictors themselves. `psych::r.test()` implements this directly (Steiger's advice, per the function's own docs), so we use that rather than a hand-rolled version.

One-tailed, same directional hypothesis as above (weighted > uniform).

```{r}
# ── Steiger-recommended Williams t-test for two dependent/overlapping correlations ──
# r12 = cor(Y, X1), r13 = cor(Y, X2), r23 = cor(X1, X2); tests r12 > r13 (one-tailed)
# psych::r.test() implements Williams (1959) as recommended by Steiger (1980).

steiger_williams_test <- function(r12, r13, r23, n) {
  res <- psych::r.test(n = n, r12 = r12, r13 = r13, r23 = r23, twotailed = FALSE)
  tibble(t_stat = res$t, df = n - 3, p_value = res$p)
}

# Pair up the two prediction sets at the same grain (concept × scale × condition)
# so we can compute r23 = cor(weighted_pred, uniform_pred).
combined_preds <- analysis_uniform %>%
  inner_join(analysis_grouped, by = c("concept", "scale_name", "condition"))

steiger_comparison <- combined_preds %>%
  group_by(concept) %>%
  summarize(
    r_weighted = cor(mean_human_rating, mean_model_prediction),
    r_uniform  = cor(mean_rating, uniform_pred),
    r_wu       = cor(mean_model_prediction, uniform_pred),
    n          = n(),
    .groups = "drop"
  ) %>%
  rowwise() %>%
  mutate(steiger_williams_test(r_weighted, r_uniform, r_wu, n)) %>%
  ungroup() %>%
  mutate(
    sig = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01  ~ "**",
      p_value < 0.05  ~ "*",
      TRUE ~ "ns"
    )
  ) %>%
  arrange(p_value)

steiger_comparison
```

### Individual Differences

Per-participant correlations between human ratings and model predictions, computed separately for the two models. The **data-aware (weighted)** model accounts for the specific maps each participant saw and the pixel-color distributions within them; the **data-agnostic (uniform)** model ignores the specific map and assumes each color in a scale was shown equally often. Both reuse the `weighted_preds` / `uniform_preds` objects built above.

#### Data-aware model: per-participant correlations

```{r}
# Per subject, at the full stimulus grain (map × scale × mapping × rotation × shift),
# join the weighted predictions, then correlate human vs predicted within each concept.
data_aware_ind_sum <- ratings_trials_exp %>%
  mutate(map_id = as.integer(map_id)) %>%
  group_by(subject_id, concept, map_id, scale_name, mapping_code, rotation_code, shift_code) %>%
  summarize(mean_rating = mean(response), .groups = "drop") %>%
  left_join(weighted_preds,
            by = c("map_id", "scale_name", "mapping_code",
                   "rotation_code", "shift_code", "concept")) %>%
  filter(!is.na(weighted_pred)) %>%
  mutate(
    mean_rating   = as.numeric(mean_rating),
    weighted_pred = as.numeric(weighted_pred)
  )

ind_corrs_aware <- data_aware_ind_sum %>%
  group_by(subject_id, concept) %>%
  summarize(
    n = n(),
    test = list(cor.test(mean_rating, weighted_pred, method = "pearson")),

    r = test[[1]]$estimate,
    p = test[[1]]$p.value,
    conf_low = test[[1]]$conf.int[1],
    conf_high = test[[1]]$conf.int[2],
    .groups = "drop"
  ) %>%
  mutate(
    sig = case_when(
      p < 0.001 ~ "***",
      p < 0.01  ~ "**",
      p < 0.05  ~ "*",
      TRUE ~ "ns"
    ),
    model = "Data-aware (weighted)"
  )
```

#### Data-agnostic model: per-participant correlations

```{r}
# Per subject, at the scale × condition grain (mirroring the aggregate uniform analysis),
# join the uniform predictions, then correlate human vs predicted within each concept.
data_agnostic_ind_sum <- ratings_trials_exp %>%
  mutate(condition = ifelse(
    (mapping_code == "scaleDef" & shift_code == "darkshift") |
    (mapping_code == "scaleRev" & shift_code == "lightshift"),
    "high_end", "low_end"
  )) %>%
  group_by(subject_id, concept, scale_name, condition) %>%
  summarize(mean_rating = mean(response), .groups = "drop") %>%
  inner_join(uniform_preds, by = c("concept", "scale_name")) %>%
  mutate(
    mean_rating  = as.numeric(mean_rating),
    uniform_pred = as.numeric(uniform_pred)
  )

ind_corrs_agnostic <- data_agnostic_ind_sum %>%
  group_by(subject_id, concept) %>%
  summarize(
    n = n(),
    test = list(cor.test(mean_rating, uniform_pred, method = "pearson")),

    r = test[[1]]$estimate,
    p = test[[1]]$p.value,
    conf_low = test[[1]]$conf.int[1],
    conf_high = test[[1]]$conf.int[2],
    .groups = "drop"
  ) %>%
  mutate(
    sig = case_when(
      p < 0.001 ~ "***",
      p < 0.01  ~ "**",
      p < 0.05  ~ "*",
      TRUE ~ "ns"
    ),
    model = "Data-agnostic (uniform)"
  )
```

#### Plotting

```{r}
# Shared histogram builder so both models render identically.
ind_corr_hist <- function(corr_df) {
  ggplot(corr_df, aes(x = r)) +
    geom_histogram(binwidth = 0.2, color = "black") +
    facet_wrap(~ concept, nrow = 1) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
    scale_x_continuous(limits = c(-.5, 1), expand = c(0, 0)) +
    labs(x = "Correlation", y = "Number of Participants") +
    theme_classic() +
    theme(
      legend.position = "none",
      panel.background = element_blank(),
      panel.grid = element_blank(),
      axis.line = element_line(color = "black"),
      panel.spacing.x = unit(2, "lines"),
      panel.spacing.y = unit(1, "lines"),
      plot.margin     = margin(10, 20, 10, 10, "pt"),
      axis.title      = element_text(size = 12),
      axis.text       = element_text(size = 12),
      strip.text      = element_text(size = 12)
    )
}

ind_plot_aware    <- ind_corr_hist(ind_corrs_aware)
ind_plot_agnostic <- ind_corr_hist(ind_corrs_agnostic)

ind_plot_aware
ind_plot_agnostic
```

```{r}
ggsave(
  paste0(plot_dir, "exp2_ind_corr_freq_dataaware.pdf"),
  ind_plot_aware,
  width = 14,
  height = 2
)

ggsave(
  paste0(plot_dir, "exp2_ind_corr_freq_dataagnostic.pdf"),
  ind_plot_agnostic,
  width = 14,
  height = 2
)
```

```{r}
tmp <- read.csv(paste0(results_dir,'matplotlib_colorscales_w_predictions_2.csv'))
tmp%>%filter(colorscale=='Greys')
```
