---
title: "Experiment 2 Analyses"
author: "Kushin Mukherjee"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(psych)
library(stringr)
library(proxy)

# Load the package
library(osfr)
```

```{r}

analysis_dir<- getwd()

data_dir <-paste0(analysis_dir,'/../data/exp_2')
### if data_dir doesn't exist create it
if (!dir.exists(data_dir)) {
  dir.create(data_dir)
}

results_dir <-paste0(analysis_dir,'/../results/')
### if results_dir doesn't exist create it
if (!dir.exists(results_dir)) {
  dir.create(results_dir)
}

plot_dir <- paste0(analysis_dir,'/../plots/')

### if plot_dir doesn't exist create it
if (!dir.exists(plot_dir)) {
  dir.create(plot_dir)
}

```

### OSF Download if needed

```{r}

osf_token <- readLines("../km_osf_pat.txt", warn = FALSE)[1]
# Authenticate with OSF (you'll be prompted to generate a PAT token)
osf_auth(token=osf_token)

# Retrieve the OSF component
osf_project <- osf_retrieve_node("6yqhd")

# List all files in the component
files <- osf_ls_files(osf_project, n_max = Inf)


# Download all files
osf_download(files, path = data_dir, conflicts = "overwrite")


```

### Load data files

```{r}
data_files <-list.files(path = data_dir, full.names = TRUE, pattern="*.csv")
print(paste0('we have data from ',length(data_files),' participants'))



df_list <- lapply(data_files, function(f) {
  df <- read.csv(f)
  df[, !names(df) %in% "SONA_id", drop = FALSE]
})

experiment_data_raw<- bind_rows(df_list)

ratings_trials<- experiment_data_raw %>% filter(trial_type== 'html-slider-response')

ratings_trials_exp<- ratings_trials %>% filter(!practiceTrial %in% c('True','true'))

### filter ratings_trials_exp such that we have only 1 subject_id per condition_num
# set.seed(56)  
# ratings_trials_exp <- ratings_trials_exp %>%
#   group_by(condition_num) %>%
#   filter(subject_id %in% sample(unique(subject_id), 1)) %>%
#   ungroup()


ratings_trials_exp$response<- as.numeric(ratings_trials_exp$response)
ratings_trials_exp$response<- (ratings_trials_exp$response + 200)/400


ratings_trials_exp_grouped <- ratings_trials_exp%>%group_by(concept,scale_name,mapping_code, shift_code)%>%summarize(mean_rating = mean(response))


### get the number of unique participants per condition in ratings_trials_exp
subject_dist<-ratings_trials_exp %>% 
  group_by(condition_num) %>% 
  summarize(num_participants = n_distinct(subject_id))

subject_dist

### can you list all the numbers between 0 and 79 not present in subject_dist$condition_num
missing_conditions <- setdiff(0:40, subject_dist$condition_num)
print(paste0('missing conditions: ', paste(missing_conditions, collapse = ', ')))



write.csv(ratings_trials_exp_grouped, '../data/exp2_ratings_trials_exp_grouped.csv')

```

```{r}
library(ggimage)

preds_per_color <- read.csv(paste0(results_dir, "matplotlib_colorscales_w_predictions_2.csv"),
                             stringsAsFactors = FALSE)
# 
# analysis_df <- ratings_trials_exp_grouped%>%inner_join(predicted_ratings_grouped_uniform_long, by = c("concept", "scale_name"))
# 
# analysis_df$path_name <- paste0("../images/small/map02_", analysis_df$scale_name, "_scaleDef_orientDef_000_25px.png")
# # 2. Build the plot
# p <- ggplot(analysis_df, aes(x = mean_prediction, y = mean_rating)) +
#   geom_image(aes(image = path_name), size = 0.05) +
#   geom_smooth(method = 'lm', se = FALSE, color = '#06c7bd') +
#   facet_wrap(~concept, nrow = 1) +
#   labs(x = 'Predicted Ratings', y = 'Human Ratings') +
#   theme_linedraw() +
#   scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
#   scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
#   coord_fixed(ratio = 1) +
#   theme(
#     panel.spacing.y = unit(1, "lines"),
#     panel.spacing.x = unit(2, "lines"),
#     plot.margin = margin(10, 20, 10, 10, "pt"),
#     axis.title = element_text(size = 12),
#     axis.text = element_text(size = 12),
#      strip.text = element_text(size = 12)
#   )
# 
# 
# # 3. Save the result
# n_facets <- length(unique(analysis_df$concept))
# panel_size <- 3.5  # inches per panel
# ggsave(
#   paste0(plot_dir, 'exp_2_scatterplot_predicted_vs_observed_ratings.pdf'),
#   plot = p,
#   width = panel_size * n_facets + 1,  # +1 for margins/labels
#   height = panel_size + 1.5,           # +1.5 for axis labels/title
#   dpi = 300
# )
# 
# p
```

### Uniform Predictions (equal-weight baseline, as in Exp1)

Assumes all 256 colors in a colorscale are present equally — the same approach used in Exp1. This serves as a baseline to compare against the proportion-weighted predictions below.

```{r}
# ── Uniform (equal-weight) predictions per colorscale ─────────────────────────

uniform_preds <- preds_per_color %>%
  group_by(colorscale) %>%
  summarize(
    angry   = mean(angry_pred),
    fearful = mean(fearful_pred),
    happy   = mean(happy_pred),
    sad     = mean(sad_pred),
    disgust = mean(disgust_pred),
    .groups = "drop"
  ) %>%
  rename(scale_name = colorscale) %>%
  pivot_longer(c(angry, fearful, happy, sad, disgust),
               names_to = "concept", values_to = "uniform_pred")

# ── Join with observed ratings and collapse to 40 points per emotion ──────────
# scaleDef+darkshift ≡ scaleRev+lightshift (both show high-end default colors)
# scaleDef+lightshift ≡ scaleRev+darkshift (both show low-end default colors)

analysis_uniform <- ratings_trials_exp_grouped %>%
  inner_join(uniform_preds, by = c("concept", "scale_name")) %>%
  mutate(condition = ifelse(
    (mapping_code == "scaleDef" & shift_code == "darkshift") |
    (mapping_code == "scaleRev" & shift_code == "lightshift"),
    "high_end", "low_end"
  )) %>%
  group_by(concept, scale_name, condition) %>%
  summarize(
    mean_rating  = mean(mean_rating),
    uniform_pred = mean(uniform_pred),
    .groups = "drop"
  ) %>%
  mutate(path_name = paste0(
    "../images/shifted_maps_small/map02_",
    scale_name, "_scaleDef_orientDef_000_",
    ifelse(condition == "high_end", "darkshift", "lightshift"), ".png"
  ))
```

```{r}
# ── Plot: uniform predicted vs observed, faceted concept × shift ──────────────

p_uniform <- ggplot(analysis_uniform, aes(x = uniform_pred, y = mean_rating)) +
  geom_image(aes(image = path_name), size = 0.06) +
  geom_smooth(method = "lm", se = FALSE, color = "#06c7bd") +
  facet_grid( cols = vars(concept)) +
  labs(x = "Uniform Predicted Rating", y = "Human Rating") +
  theme_linedraw() +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
  coord_fixed(ratio = 1) +
  theme(
    panel.spacing.x = unit(2, "lines"),
    panel.spacing.y = unit(1, "lines"),
    plot.margin     = margin(10, 20, 10, 10, "pt"),
    axis.title      = element_text(size = 12),
    axis.text       = element_text(size = 12),
    strip.text      = element_text(size = 12)
  )

n_concepts <- length(unique(analysis_uniform$concept))
panel_size <- 3.5

ggsave(
  paste0(plot_dir, "exp_2_scatterplot_uniform_pred_vs_observed.pdf"),
  plot   = p_uniform,
  width  = panel_size * n_concepts + 1,
  height = panel_size * 2 + 2,
  dpi    = 300
)

p_uniform
```

```{r}
# ── Correlation: uniform predictions vs observed (per concept) ────────────────

analysis_uniform %>%
  group_by(concept) %>%
  summarize(
    r    = cor(uniform_pred, mean_rating, use = "complete.obs"),
    n    = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(r))
```

### Weighted Predictions (color-proportion-weighted)

Exp1 predictions assumed all 256 colors in a scale appear equally. Here we weight each color's predicted affect rating by the proportion of map pixels that actually display that color. Proportions were pre-computed per (map_id × scale_name × mapping_code) by `prep/compute_color_proportions.R` and are orientation/rotation/shift invariant (same pixels, different arrangement or brightness). The join to human data is at the full stimulus grain: map × scale × mapping × rotation × shift (3200 unique stimuli), with weighted_pred broadcast across rotation and shift variants.

```{r}
# ── Load proportions ──────────────────────────────────────────────────────────

proportions <- read.csv(paste0(results_dir, "color_proportions.csv"),
                        stringsAsFactors = FALSE)

# ── Pivot predictions to long format ──────────────────────────────────────────

concept_cols <- c("angry_pred", "happy_pred", "fearful_pred", "disgust_pred", "sad_pred")

preds_long <- preds_per_color %>%
  select(colorscale, index, all_of(concept_cols)) %>%
  pivot_longer(all_of(concept_cols), names_to = "concept", values_to = "pred") %>%
  mutate(concept = str_remove(concept, "_pred"))

# ── Compute weighted predictions ───────────────────────────────────────────────
# proportion × pred, summed over index, per stimulus × concept

weighted_preds <- proportions %>%
  left_join(preds_long, by = c("scale_name" = "colorscale", "index")) %>%
  group_by(map_id, scale_name, mapping_code, rotation_code, shift_code, concept) %>%
  summarize(weighted_pred = sum(proportion * pred), .groups = "drop")

weighted_preds
```

```{r}
# ── Fine-grained observed ratings ─────────────────────────────────────────────
# Group at the unique-stimulus level: map × scale × mapping × rotation × shift.
# Orientation is collapsed (mirroring preserves color proportions and counts).
# rotation_code in the data uses bare integers ("0","90",...); filenames use
# zero-padded strings ("000","090",...) — conversion handled in path construction.

ratings_fine <- ratings_trials_exp %>%
  mutate(map_id = as.integer(map_id)) %>%
  group_by(concept, map_id, scale_name, mapping_code, rotation_code, shift_code) %>%
  summarize(mean_rating = mean(response), .groups = "drop")

# ── Join observed ratings with weighted predictions ────────────────────────────

analysis_weighted <- ratings_fine %>%
  left_join(weighted_preds,
            by = c("map_id", "scale_name", "mapping_code",
                   "rotation_code", "shift_code", "concept"))

analysis_weighted
```

```{r}
# ── Plot: weighted predicted vs observed ratings, faceted by concept ───────────

# Collapse to 40 points per emotion (20 scales × 2 conditions)
# scaleDef+darkshift ≡ scaleRev+lightshift (both show high-end default colors)
# scaleDef+lightshift ≡ scaleRev+darkshift (both show low-end default colors)
analysis_grouped <- analysis_weighted %>%
  mutate(condition = ifelse(
    (mapping_code == "scaleDef" & shift_code == "darkshift") |
    (mapping_code == "scaleRev" & shift_code == "lightshift"),
    "high_end", "low_end"
  )) %>%
  group_by(concept, scale_name, condition) %>%
  summarize(
    mean_human_rating     = mean(mean_rating),
    mean_model_prediction = mean(weighted_pred),
    .groups = "drop"
  ) %>%
  mutate(path_name = paste0(
    "../images/shifted_maps_small/map02_",
    scale_name, "_scaleDef_orientDef_000_",
    ifelse(condition == "high_end", "darkshift", "lightshift"), ".png"
  ))

p_weighted <- ggplot(analysis_grouped, aes(x = mean_model_prediction, y = mean_human_rating)) +
  geom_image(aes(image = path_name), size = 0.06) +
  geom_smooth(method = "lm", se = FALSE, color = "#06c7bd") +
  facet_grid( cols = vars(concept)) +
  labs(x = "Weighted Predicted Rating", y = "Human Rating") +
  theme_linedraw() +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
  coord_fixed(ratio = 1) +
  theme(
    panel.spacing.x  = unit(2, "lines"),
    panel.spacing.y  = unit(1, "lines"),
    plot.margin      = margin(10, 20, 10, 10, "pt"),
    axis.title       = element_text(size = 12),
    axis.text        = element_text(size = 12),
    strip.text       = element_text(size = 12)
  )

n_concepts <- length(unique(analysis_grouped$concept))
panel_size <- 3.5

ggsave(
  paste0(plot_dir, "exp_2_scatterplot_weighted_pred_vs_observed.pdf"),
  plot   = p_weighted,
  width  = panel_size * n_concepts + 1,
  height = panel_size * 2 + 2,
  dpi    = 300
)

p_weighted
```

```{r}
# ── Correlation: weighted predictions vs observed (per concept) ────────────────

# analysis_weighted %>%
#   group_by(concept) %>%
#   summarize(
#     r     = cor(weighted_pred, mean_rating, use = "complete.obs"),
#     r_sq  = r^2,
#     n     = n(),
#     .groups = "drop"
#   ) %>%
#   arrange(desc(r))


analysis_grouped %>%
  group_by(concept) %>%
  summarize(
    r     = cor(mean_human_rating, mean_model_prediction, use = "complete.obs"),
    n     = n(),
  
    .groups = "drop"
  ) %>%
  arrange(desc(r))
```

```{r}
tmp <- read.csv(paste0(results_dir,'matplotlib_colorscales_w_predictions_2.csv'))
tmp%>%filter(colorscale=='Greys')
```
