---
title: "exp3_analysis"
author: "Anna Chinni"
format: pdf
editor: visual
---

Analysis script for experiment 3 

```{r}
rm(list = ls())
```

```{r}
# List of packages
packages <- c("corrr", "ggplot2", "lme4", "lmerTest", "FactoMineR", "grid","png","cowplot","magick", "purrr", "combinat", "ggimage","dplyr","tidyverse","knitr", "readxl", "writexl", "tidytext", "forcats", "ggrepel", "stringr", "jsonlite", "MuMIn")
```

```{r}

install_and_load_packages <- function(packages) {
  for(package in packages) {
    if(!require(package, character.only = TRUE)) {
      install.packages(package, dependencies = TRUE)
      library(package, character.only = TRUE)
    }
  }
}

# Call the function with the list of packages
install_and_load_packages(packages)

```

```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

```

```{r}
current_dir <- getwd()

dots <- read_csv("../data/exp_3/dots_data_clean.csv")

bars <- read_csv("../data/exp_3/bars_data_clean.csv")

StimProp <- read.csv("../data/exp_3/StimProp.csv")

```

```{r}
dots <- dots%>%
  left_join(StimProp, by=c("SubjNum", "Emotion", "AssocCond", "Version", "Palette", "ColorOrder", "YOrder"))

bars <- bars%>%
  left_join(StimProp, by=c("SubjNum", "Emotion", "AssocCond", "Version", "Palette", "ColorOrder", "YOrder"))
```

```{r}
data <- rbind(dots, bars)
```

```{r}
#code whether strong associate colors were higher Y-axis value or weaker were higher Y-values
data <- data %>% 
  mutate(
    arrange = case_when(
      YOrder == "lgsm" & ColorOrder == "decend" ~ "strong_big",
      YOrder == "smlg" & ColorOrder == "ascend" ~ "strong_big", 
      YOrder == "lgsm" & ColorOrder == "ascend" ~ "weak_big",
      YOrder == "smlg" & ColorOrder == "decend" ~ "weak_big"
    )
  )

```

```{r}
#created weigthed association for each color (Each Emo is the predicted association for a color with a given emotion)
data <- data%>%
  mutate(
    YEmo1 = Yval1*Emo1,
    YEmo2 = Yval2*Emo2,
    YEmo3 = Yval3*Emo3,
    YEmo4 = Yval4*Emo4
  )

```

```{r}
#created average assocaition score and weighted average association score
data <- data%>%
  mutate(
    AvgEmoPred=rowMeans(select(.,Emo1, Emo2, Emo3, Emo4)),
    YAvgEmoPred=rowMeans(select(.,YEmo1, YEmo2, YEmo3, YEmo4))
  )
```

```{r}
#scale YAvgEmoPred so it ranges from 0 to 1
data <- data %>% 
  mutate(
    unscaled_YAvgEmoPre = YAvgEmoPred,
    YAvgEmoPred = (YAvgEmoPred - min(YAvgEmoPred)) / 
                          (max(YAvgEmoPred) - min(YAvgEmoPred))
  )
```

# Weigthed versus Unweighted Association Correlations

## Data Prep

```{r}
#summarize data for plotting
data_sum <- data %>% 
  group_by(plot, concept, AssocCond, Version, Palette, arrange) %>% 
  summarise(
    AvgEmoPred = mean(AvgEmoPred),
    YAvgEmoPred = mean(YAvgEmoPred),
    mean_rating = mean(response)
  ) %>% 
  ungroup()
```

## Correlations

```{r}
#correlations for weighted (wa) and unweighted (a) ratings based on Y-axis values.
variable_corr <- list(
  
  #Angry
  "dots_angry_a"   = cor.test(data_sum %>% filter(plot == "dot" & concept == "Angry") %>% pull(mean_rating), data_sum %>% filter(plot == "dot" & concept == "Angry") %>% pull(AvgEmoPred)),
  "dots_angry_wa"  = cor.test(data_sum %>% filter(plot == "dot" & concept == "Angry") %>% pull(mean_rating), data_sum %>% filter(plot == "dot" & concept == "Angry") %>% pull(YAvgEmoPred)),
  "bars_angry_a"   = cor.test(data_sum %>% filter(plot == "bar" & concept == "Angry") %>% pull(mean_rating), data_sum %>% filter(plot == "bar" & concept == "Angry") %>% pull(AvgEmoPred)),
  "bars_angry_wa"  = cor.test(data_sum %>% filter(plot == "bar" & concept == "Angry") %>% pull(mean_rating), data_sum %>% filter(plot == "bar" & concept == "Angry") %>% pull(YAvgEmoPred)),
  
  #Disgust
  "dots_disgust_a" = cor.test(data_sum %>% filter(plot == "dot" & concept == "Disgust") %>% pull(mean_rating), data_sum %>% filter(plot == "dot" & concept == "Disgust") %>% pull(AvgEmoPred)),
  "dots_disgust_wa"= cor.test(data_sum %>% filter(plot == "dot" & concept == "Disgust") %>% pull(mean_rating), data_sum %>% filter(plot == "dot" & concept == "Disgust") %>% pull(YAvgEmoPred)),
  "bars_disgust_a" = cor.test(data_sum %>% filter(plot == "bar" & concept == "Disgust") %>% pull(mean_rating), data_sum %>% filter(plot == "bar" & concept == "Disgust") %>% pull(AvgEmoPred)),
  "bars_disgust_wa"= cor.test(data_sum %>% filter(plot == "bar" & concept == "Disgust") %>% pull(mean_rating), data_sum %>% filter(plot == "bar" & concept == "Disgust") %>% pull(YAvgEmoPred)),
  
  #Fearful
  "dots_fearful_a" = cor.test(data_sum %>% filter(plot == "dot" & concept == "Fearful") %>% pull(mean_rating), data_sum %>% filter(plot == "dot" & concept == "Fearful") %>% pull(AvgEmoPred)),
  "dots_fearful_wa"= cor.test(data_sum %>% filter(plot == "dot" & concept == "Fearful") %>% pull(mean_rating), data_sum %>% filter(plot == "dot" & concept == "Fearful") %>% pull(YAvgEmoPred)),
  "bars_fearful_a" = cor.test(data_sum %>% filter(plot == "bar" & concept == "Fearful") %>% pull(mean_rating), data_sum %>% filter(plot == "bar" & concept == "Fearful") %>% pull(AvgEmoPred)),
  "bars_fearful_wa"= cor.test(data_sum %>% filter(plot == "bar" & concept == "Fearful") %>% pull(mean_rating), data_sum %>% filter(plot == "bar" & concept == "Fearful") %>% pull(YAvgEmoPred)),
  
  #Happy
  "dots_happy_a"   = cor.test(data_sum %>% filter(plot == "dot" & concept == "Happy") %>% pull(mean_rating), data_sum %>% filter(plot == "dot" & concept == "Happy") %>% pull(AvgEmoPred)),
  "dots_happy_wa"  = cor.test(data_sum %>% filter(plot == "dot" & concept == "Happy") %>% pull(mean_rating), data_sum %>% filter(plot == "dot" & concept == "Happy") %>% pull(YAvgEmoPred)),
  "bars_happy_a"   = cor.test(data_sum %>% filter(plot == "bar" & concept == "Happy") %>% pull(mean_rating), data_sum %>% filter(plot == "bar" & concept == "Happy") %>% pull(AvgEmoPred)),
  "bars_happy_wa"  = cor.test(data_sum %>% filter(plot == "bar" & concept == "Happy") %>% pull(mean_rating), data_sum %>% filter(plot == "bar" & concept == "Happy") %>% pull(YAvgEmoPred)),
  
  #Sad
  "dots_sad_a"     = cor.test(data_sum %>% filter(plot == "dot" & concept == "Sad") %>% pull(mean_rating), data_sum %>% filter(plot == "dot" & concept == "Sad") %>% pull(AvgEmoPred)),
  "dots_sad_wa"    = cor.test(data_sum %>% filter(plot == "dot" & concept == "Sad") %>% pull(mean_rating), data_sum %>% filter(plot == "dot" & concept == "Sad") %>% pull(YAvgEmoPred)),
  "bars_sad_a"     = cor.test(data_sum %>% filter(plot == "bar" & concept == "Sad") %>% pull(mean_rating), data_sum %>% filter(plot == "bar" & concept == "Sad") %>% pull(AvgEmoPred)),
  "bars_sad_wa"    = cor.test(data_sum %>% filter(plot == "bar" & concept == "Sad") %>% pull(mean_rating), data_sum %>% filter(plot == "bar" & concept == "Sad") %>% pull(YAvgEmoPred))
)

```

```{r}
#create table of correlations
r_var_table <- tibble(
  Model = names(variable_corr),
  R = sapply(variable_corr, function(x) x$estimate),
  t = sapply(variable_corr, function(x) x$statistic),
  df = sapply(variable_corr, function(x) x$parameter[1]),
  p = sapply(variable_corr, function(x) x$p.value),
)
```

## Plotting

```{r}
#add file path for squares
data_sum <- data_sum %>% 
  mutate(
    filepath = paste0(
      current_dir,"../images/dots_bars_small/",
      concept, "_", AssocCond, "_", Version, "_", Palette, "_", arrange,
      ".png"
    )
  )
```

```{r}
data_sum$arrange <- as_factor(data_sum$arrange)
```

```{r}
#grab values
r_var_table <- r_var_table %>% 
  mutate(
    plot  = word(Model, 1, sep = "_"),
    concept    = str_to_title(word(Model, 2, sep = "_")),
    model_type = word(Model, 3, sep = "_")
  )

#add labels info
r_labels <- r_var_table %>%
  mutate(
    label = paste0("R = ", round(R, 2), ", p =", round(p, 3)),
    x = 0.95,    # x-position in panel (near right)
    y = 0.05     # y-position in panel (near bottom)
  )
```

```{r}
p_assoc_dots <- ggplot(data_sum %>% filter(plot == "dot"), aes(x = AvgEmoPred, y = mean_rating)) +
#   geom_point(
#   aes(color = arrange),
#   shape = 15,   # square
#   size = 8
# )+
  geom_image(aes(image = filepath), size = 0.06) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  facet_grid( cols = vars(concept)) +
  labs(x = "Dots: Predicted Association Rating", y = "Human Rating") +
#   scale_color_manual(
#   values = c(
#     "strong_big" = "black",
#     "weak_big" = "lightgrey"
#   )
# ) +
  theme_linedraw() +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
  coord_fixed(ratio = 1) +
  theme(
    legend.position = "none",
    panel.spacing.x = unit(2, "lines"),
    panel.spacing.y = unit(1, "lines"),
    plot.margin     = margin(10, 20, 10, 10, "pt"),
    axis.title      = element_text(size = 12),
    axis.text       = element_text(size = 12),
    strip.text      = element_text(size = 12)
  ) +
  geom_text(
    data = r_labels %>% filter(plot == "dots" & model_type == "a"),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,   # important so it doesn't inherit main plot aesthetics
    hjust = 1,             # right-align
    vjust = 0,             # bottom-align
    size = 4,
    color = "black"
  )


p_assoc_dots

```

```{r}
p_assoc_bars <- ggplot(data_sum %>% filter(plot == "bar"), aes(x = AvgEmoPred, y = mean_rating)) +
#   geom_point(
#   aes(color = arrange),
#   shape = 15,   # square
#   size = 8
# )+
  geom_image(aes(image = filepath), size = 0.06) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  facet_grid( cols = vars(concept)) +
  labs(x = "Bars: Predicted Association Rating", y = "Human Rating") +
#   scale_color_manual(
#   values = c(
#     "strong_big" = "black",
#     "weak_big" = "lightgrey"
#   )
# ) +
  theme_linedraw() +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0)) +
  coord_fixed(ratio = 1) +
  theme(
    legend.position = "none",
    panel.spacing.x = unit(2, "lines"),
    panel.spacing.y = unit(1, "lines"),
    plot.margin     = margin(10, 20, 10, 10, "pt"),
    axis.title      = element_text(size = 12),
    axis.text       = element_text(size = 12),
    strip.text      = element_text(size = 12)
  ) +
  geom_text(
    data = r_labels %>% filter(plot == "bars" & model_type == "a"),
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,   # important so it doesn't inherit main plot aesthetics
    hjust = 1,             # right-align
    vjust = 0,             # bottom-align
    size = 4,
    color = "black"
  )


p_assoc_bars

```

```{r}
#Save the result
n_facets <- length(unique(bars$concept))
panel_size <- 3.5  # inches per panel
ggsave(
  filename = 'exp_3_scatterplot_predicted_vs_observed_ratings_dots.pdf', 
  plot = p_assoc_dots, 
  width = panel_size * n_facets + 1,  # +1 for margins/labels
  height = panel_size + 1.5,           # +1.5 for axis labels/title
  dpi = 300
)
```

```{r}
#Save the result
n_facets <- length(unique(bars$concept))
panel_size <- 3.5  # inches per panel
ggsave(
  filename = 'exp_3_scatterplot_predicted_vs_observed_ratings_bars.pdf', 
  plot = p_assoc_bars, 
  width = panel_size * n_facets + 1,  # +1 for margins/labels
  height = panel_size + 1.5,           # +1.5 for axis labels/title
  dpi = 300
)
```

#Difference Scores

```{r}
#pivot data arrange values are separate columns to compare across same plots with different y-values for same level of associated colors
diff_data <- data %>%
  select(subject_id, plot, AssocCond, Version, Palette, concept, arrange, response) %>%
  pivot_wider(
    names_from = arrange,
    values_from = response
    ) 

#the pivot created a list as we had two of the same plots (except the order of the colors with each height), so unlist and average across. Then calculate difference score. 
diff_data <- diff_data %>%
  mutate(
    strong_big = map_dbl(strong_big, mean),
    weak_big   = map_dbl(weak_big, mean),
    diff_score = strong_big - weak_big
  )

```

```{r}
#do the same for unweighted predictions - just to get our new Avg predictor, there will be no difference in this case.
pred_diff_data <- data %>% 
  select(subject_id, plot, AssocCond, Version, Palette, concept, arrange, AvgEmoPred) %>%
  pivot_wider(
    names_from = arrange,
    values_from = AvgEmoPred
    ) 

pred_diff_data <- pred_diff_data %>%
  mutate(
    pred_strong_big = map_dbl(strong_big, mean),
    pred_weak_big   = map_dbl(weak_big, mean)
  )

#recalc AvgEmoPred
pred_diff_data <- pred_diff_data %>%
  mutate(
    AvgEmoPred = rowMeans(select(.,pred_strong_big,pred_weak_big))
  )

```

```{r}
#same for weighted predictions
  
Ypred_diff_data <- data %>% 
  select(subject_id, plot, AssocCond, Version, Palette, concept, arrange, YAvgEmoPred) %>%
  pivot_wider(
    names_from = arrange,
    values_from = YAvgEmoPred
    ) 

Ypred_diff_data <- Ypred_diff_data %>%
  mutate(
    Ypred_strong_big = map_dbl(strong_big, mean),
    Ypred_weak_big   = map_dbl(weak_big, mean),
    Ypred_diff_score = Ypred_strong_big - Ypred_weak_big
  )

#recalc YAvgEmoPred
Ypred_diff_data <- Ypred_diff_data %>%
  mutate(
    YAvgEmoPred = rowMeans(select(.,Ypred_strong_big,Ypred_weak_big))
  )

```

```{r}
#combine data

diff_data <- diff_data %>% 
  left_join(pred_diff_data %>% select(subject_id, plot, AssocCond, Version, Palette, concept, pred_strong_big, pred_weak_big, AvgEmoPred), by = c("subject_id", "plot", "AssocCond", "Version", "Palette", "concept"))


diff_data <- diff_data %>% 
  left_join(Ypred_diff_data %>% select(subject_id, plot, AssocCond, Version, Palette, concept, Ypred_strong_big, Ypred_weak_big, Ypred_diff_score, YAvgEmoPred), by = c("subject_id", "plot", "AssocCond", "Version", "Palette", "concept"))
```

```{r}
#get difference in predicted association for each palette and AssocCond

diff_data <- diff_data %>% 
  left_join(data %>% select(plot, AssocCond, Version, Palette, Emo1, Emo2, Emo3, Emo4), by = c("plot", "AssocCond", "Version", "Palette"))

diff_data <- diff_data %>% 
  mutate(
    assoc_1_2 = rowMeans(select(., Emo1, Emo2)),
    assoc_3_4 = rowMeans(select(., Emo3, Emo4))
  )

#calculate the difference in association stregth between stronger and weaker association - use abs() as the strong assocaites could be Emo 1 and 2 or Emo 3 and 4
diff_data <- diff_data %>% 
  mutate(
    assoc_diff = abs(assoc_1_2 - assoc_3_4)
  )


```

## Correlations

```{r}
diff_data_sum <- diff_data %>% 
  group_by(plot, concept, AssocCond, Version, Palette) %>% 
  summarise(
    mean_diff = mean(diff_score),
    mean_AvgEmoPred = mean(AvgEmoPred),
    mean_YAvgEmoPred = mean(YAvgEmoPred),
    mean_YAvgEmoPred_diff = mean(Ypred_diff_score),
    mean_assoc_diff = mean(assoc_diff)
  ) %>% 
  ungroup()

```

```{r}
diff_corr <- list(
  
  #Angry
  "dots_angry_wa"  = cor.test(diff_data_sum %>% filter(plot == "dot" & concept == "Angry") %>% pull(mean_diff), diff_data_sum %>% filter(plot == "dot" & concept == "Angry") %>% pull(mean_YAvgEmoPred_diff)),
  "bars_angry_wa"  = cor.test(diff_data_sum %>% filter(plot == "bar" & concept == "Angry") %>% pull(mean_diff), diff_data_sum %>% filter(plot == "bar" & concept == "Angry") %>% pull(mean_YAvgEmoPred_diff)),
  
  #Disgust
  "dots_disgust_wa"= cor.test(diff_data_sum %>% filter(plot == "dot" & concept == "Disgust") %>% pull(mean_diff), diff_data_sum %>% filter(plot == "dot" & concept == "Disgust") %>% pull(mean_YAvgEmoPred_diff)),
  "bars_disgust_wa"= cor.test(diff_data_sum %>% filter(plot == "bar" & concept == "Disgust") %>% pull(mean_diff), diff_data_sum %>% filter(plot == "bar" & concept == "Disgust") %>% pull(mean_YAvgEmoPred_diff)),
  
  #Fearful
  "dots_fearful_wa"= cor.test(diff_data_sum %>% filter(plot == "dot" & concept == "Fearful") %>% pull(mean_diff), diff_data_sum %>% filter(plot == "dot" & concept == "Fearful") %>% pull(mean_YAvgEmoPred_diff)),
  "bars_fearful_wa"= cor.test(diff_data_sum %>% filter(plot == "bar" & concept == "Fearful") %>% pull(mean_diff), diff_data_sum %>% filter(plot == "bar" & concept == "Fearful") %>% pull(mean_YAvgEmoPred_diff)),
  
  #Happy
  "dots_happy_wa"  = cor.test(diff_data_sum %>% filter(plot == "dot" & concept == "Happy") %>% pull(mean_diff), diff_data_sum %>% filter(plot == "dot" & concept == "Happy") %>% pull(mean_YAvgEmoPred_diff)),
  "bars_happy_wa"  = cor.test(diff_data_sum %>% filter(plot == "bar" & concept == "Happy") %>% pull(mean_diff), diff_data_sum %>% filter(plot == "bar" & concept == "Happy") %>% pull(mean_YAvgEmoPred_diff)),
  
  #Sad
  "dots_sad_wa"    = cor.test(diff_data_sum %>% filter(plot == "dot" & concept == "Sad") %>% pull(mean_diff), diff_data_sum %>% filter(plot == "dot" & concept == "Sad") %>% pull(mean_YAvgEmoPred_diff)),
  "bars_sad_wa"    = cor.test(diff_data_sum %>% filter(plot == "bar" & concept == "Sad") %>% pull(mean_diff), diff_data_sum %>% filter(plot == "bar" & concept == "Sad") %>% pull(mean_YAvgEmoPred_diff))
)

```

```{r}
r_diff_table <- tibble(
  Model = names(diff_corr),
  R = sapply(diff_corr, function(x) x$estimate),
  t = sapply(diff_corr, function(x) x$statistic),
  df = sapply(diff_corr, function(x) x$parameter[1]),
  p = sapply(diff_corr, function(x) x$p.value),
)
```

```{r}
r_diff_table <- r_diff_table %>% 
  mutate(
    plot  = word(Model, 1, sep = "_"),
    concept    = str_to_title(word(Model, 2, sep = "_")),
    model_type = word(Model, 3, sep = "_")
  )

r_diff_labels <- r_diff_table %>%
  mutate(
    label = paste0("R = ", round(R, 2), ", p = ", round(p, 3)),
    x = 0.45,    # x-position in panel (near right)
    y = -0.2     # y-position in panel (near bottom)
  )
```

```{r}
assoc_diff_corr <- list(
  
  #Angry
  "dots_angry_a"  = cor.test(diff_data_sum %>% filter(plot == "dot" & concept == "Angry") %>% pull(mean_diff), diff_data_sum %>% filter(plot == "dot" & concept == "Angry") %>% pull(mean_assoc_diff)),
  "bars_angry_a"  = cor.test(diff_data_sum %>% filter(plot == "bar" & concept == "Angry") %>% pull(mean_diff), diff_data_sum %>% filter(plot == "bar" & concept == "Angry") %>% pull(mean_assoc_diff)),
  
  #Disgust
  "dots_disgust_a"= cor.test(diff_data_sum %>% filter(plot == "dot" & concept == "Disgust") %>% pull(mean_diff), diff_data_sum %>% filter(plot == "dot" & concept == "Disgust") %>% pull(mean_assoc_diff)),
  "bars_disgust_a"= cor.test(diff_data_sum %>% filter(plot == "bar" & concept == "Disgust") %>% pull(mean_diff), diff_data_sum %>% filter(plot == "bar" & concept == "Disgust") %>% pull(mean_assoc_diff)),
  
  #Fearful
  "dots_fearful_a"= cor.test(diff_data_sum %>% filter(plot == "dot" & concept == "Fearful") %>% pull(mean_diff), diff_data_sum %>% filter(plot == "dot" & concept == "Fearful") %>% pull(mean_assoc_diff)),
  "bars_fearful_a"= cor.test(diff_data_sum %>% filter(plot == "bar" & concept == "Fearful") %>% pull(mean_diff), diff_data_sum %>% filter(plot == "bar" & concept == "Fearful") %>% pull(mean_assoc_diff)),
  
  #Happy
  "dots_happy_a"  = cor.test(diff_data_sum %>% filter(plot == "dot" & concept == "Happy") %>% pull(mean_diff), diff_data_sum %>% filter(plot == "dot" & concept == "Happy") %>% pull(mean_assoc_diff)),
  "bars_happy_a"  = cor.test(diff_data_sum %>% filter(plot == "bar" & concept == "Happy") %>% pull(mean_diff), diff_data_sum %>% filter(plot == "bar" & concept == "Happy") %>% pull(mean_assoc_diff)),
  
  #Sad
  "dots_sad_a"    = cor.test(diff_data_sum %>% filter(plot == "dot" & concept == "Sad") %>% pull(mean_diff), diff_data_sum %>% filter(plot == "dot" & concept == "Sad") %>% pull(mean_assoc_diff)),
  "bars_sad_a"    = cor.test(diff_data_sum %>% filter(plot == "bar" & concept == "Sad") %>% pull(mean_diff), diff_data_sum %>% filter(plot == "bar" & concept == "Sad") %>% pull(mean_assoc_diff))
)

```

```{r}
r_assoc_diff_table <- tibble(
  Model = names(assoc_diff_corr),
  R = sapply(assoc_diff_corr, function(x) x$estimate),
  t = sapply(assoc_diff_corr, function(x) x$statistic),
  df = sapply(assoc_diff_corr, function(x) x$parameter[1]),
  p = sapply(assoc_diff_corr, function(x) x$p.value),
)
```

```{r}
r_assoc_diff_table <- r_assoc_diff_table %>% 
  mutate(
    plot  = word(Model, 1, sep = "_"),
    concept    = str_to_title(word(Model, 2, sep = "_")),
    model_type = word(Model, 3, sep = "_")
  )

r_assoc_diff_labels <- r_assoc_diff_table %>%
  mutate(
    label = paste0("R = ", round(R, 2), ", p = ", round(p, 3)),
    x = 0.45,    # x-position in panel (near right)
    y = -0.2     # y-position in panel (near bottom)
  )
```

```{r}
#add file path for squares
diff_data_sum <- diff_data_sum %>% 
  mutate(
    filepath = paste0(
      current_dir,"../images/dots_bars_small/",
      concept, "_", AssocCond, "_", Version, "_", Palette,
      ".png"
    )
  )
```

# Plot Dots and Bars Assoc Diff Together

Plots of predicted association difference for the stronger and weaker assocaited colors in each plot

```{r}
p_assoc_diff <- ggplot(diff_data_sum, aes(x = mean_assoc_diff, y = mean_diff)) +
  #geom_image(aes(image = filepath), size = 0.06) +
  geom_point(
  aes(color = plot,
      shape = plot),
  size = 4,
  alpha = 0.75 
  )+
  geom_smooth(aes(group = plot, color = plot), method = "lm", se = FALSE) +
  facet_grid( cols = vars(concept)) +
  labs(x = "Predicted Association Rating Difference Score", y = "Human rating Difference Score") +
  scale_shape_manual(values = c("dot" = 16, "bar" = 15)) +
  scale_color_manual(
  values = c(
    "bar" = "#636363",
    "dot" = "#bdbdbd"
  )
  ) +
  theme_linedraw() +
  scale_y_continuous(limits = c(-.1, .3), expand = c(0, 0)) +
  scale_x_continuous(limits = c(0, .6), expand = c(0, 0)) +
  coord_fixed(ratio = 1) +
  theme(
    legend.position = "none",
    panel.spacing.x = unit(2, "lines"),
    panel.spacing.y = unit(1, "lines"),
    plot.margin     = margin(10, 20, 10, 10, "pt"),
    axis.title      = element_text(size = 12),
    axis.text       = element_text(size = 12),
    strip.text      = element_text(size = 12)
  )

p_assoc_diff
```

```{r}
#Save the result
n_facets <- length(unique(bars$concept))
panel_size <- 3.5  # inches per panel
ggsave(
  filename = 'exp_3_assoc_diff_scores.pdf', 
  plot = p_assoc_diff, 
  width = panel_size * n_facets + 1,  # +1 for margins/labels
  height = panel_size + 1.5,           # +1.5 for axis labels/title
  dpi = 300
)
```

# Models

Models include separate models for each emotion All - the association difference with plort type full interactive model Dots - assocaition difference for dots Bars - association difference for bars

```{r}
#center predictor for interaction model

diff_data_int <- diff_data %>% 
  group_by(concept) %>% 
  mutate(
    Ypred_diff_score_c = Ypred_diff_score - mean(Ypred_diff_score),
    assoc_diff_c = assoc_diff - mean(assoc_diff),
    plot_c = if_else(plot == "bar", .5, -.5)
  ) %>% 
  ungroup()
```

```{r}
#filter by emotion
angry_diff <- diff_data_int %>% 
  filter(concept == "Angry")

disgust_diff <- diff_data_int %>% 
  filter(concept == "Disgust")

fearful_diff <- diff_data_int %>% 
  filter(concept == "Fearful")

happy_diff <- diff_data_int %>% 
  filter(concept == "Happy")

sad_diff <- diff_data_int %>% 
  filter(concept == "Sad")
```

```{r}
#center predictor for plot specific models

diff_data_plots <- diff_data %>% 
  group_by(plot, concept) %>% 
  mutate(
    Ypred_diff_score_c = Ypred_diff_score - mean(Ypred_diff_score),
    assoc_diff_c = assoc_diff - mean(assoc_diff),
    plot_c = if_else(plot == "bar", .5, -.5)
  ) %>% 
  ungroup()
```

```{r}
#filter by plot and emotion
angry_diff_dots <- diff_data_plots %>% 
  filter(plot == "dot" & concept == "Angry")
angry_diff_bars <- diff_data_plots %>% 
  filter(plot == "bar" & concept == "Angry")

disgust_diff_dots <- diff_data_plots %>% 
  filter(plot == "dot" & concept == "Disgust")
disgust_diff_bars <- diff_data_plots %>% 
  filter(plot == "bar" & concept == "Disgust")

fearful_diff_dots <- diff_data_plots %>% 
  filter(plot == "dot" & concept == "Fearful")
fearful_diff_bars <- diff_data_plots %>% 
  filter(plot == "bar" & concept == "Fearful")

happy_diff_dots <- diff_data_plots %>% 
  filter(plot == "dot" & concept == "Happy")
happy_diff_bars <- diff_data_plots %>% 
  filter(plot == "bar" & concept == "Happy")

sad_diff_dots <- diff_data_plots %>% 
  filter(plot == "dot" & concept == "Sad")
sad_diff_bars <- diff_data_plots %>% 
  filter(plot == "bar" & concept == "Sad")
```

## Angry

### All

```{r}
angry_m <- lmer(diff_score~assoc_diff_c*plot_c+(1+assoc_diff_c|subject_id),
                data=angry_diff,
                verbose=F,
                control=lmerControl(
                  optimizer="bobyqa",
                  optCtrl=list(maxfun=2e5)
                ))
summary(angry_m)
```

### Dots

```{r}
angry_dots_m <- lmer(diff_score~assoc_diff_c+(1+assoc_diff_c|subject_id),
                data=angry_diff_dots,
                verbose=F,
                control=lmerControl(
                  optimizer="bobyqa",
                  optCtrl=list(maxfun=2e5)
                ))
summary(angry_dots_m)
```

### Bars

```{r}
angry_bars_m <- lmer(diff_score~assoc_diff_c+(1+assoc_diff_c|subject_id),
                data=angry_diff_bars,
                verbose=F,
                control=lmerControl(
                  optimizer="bobyqa",
                  optCtrl=list(maxfun=2e5)
                ))
summary(angry_bars_m)
```

## Disgust

### All

```{r}
disgust_m <- lmer(diff_score~assoc_diff_c*plot_c+(1+assoc_diff_c|subject_id),
                data=disgust_diff,
                verbose=F,
                control=lmerControl(
                  optimizer="bobyqa",
                  optCtrl=list(maxfun=2e5)
                ))
summary(disgust_m)
```

### Dots

```{r}
disgust_dots_m <- lmer(diff_score~assoc_diff_c+(1+assoc_diff_c|subject_id),
                data=disgust_diff_dots,
                verbose=F,
                control=lmerControl(
                  optimizer="bobyqa",
                  optCtrl=list(maxfun=2e5)
                ))
summary(disgust_dots_m)
```

### Bars

```{r}
disgust_bars_m <- lmer(diff_score~assoc_diff_c+(1+assoc_diff_c|subject_id),
                data=disgust_diff_bars,
                verbose=F,
                control=lmerControl(
                  optimizer="bobyqa",
                  optCtrl=list(maxfun=2e5)
                ))
summary(disgust_bars_m)
```

## Fearful

### All

```{r}
fearful_m <- lmer(diff_score~assoc_diff_c*plot_c+(1+assoc_diff_c|subject_id),
                data=fearful_diff,
                verbose=F,
                control=lmerControl(
                  optimizer="bobyqa",
                  optCtrl=list(maxfun=2e5)
                ))
summary(fearful_m)
```

### Dots

```{r}
fearful_dots_m <- lmer(diff_score~assoc_diff_c+(1+assoc_diff_c|subject_id),
                data=fearful_diff_dots,
                verbose=F,
                control=lmerControl(
                  optimizer="bobyqa",
                  optCtrl=list(maxfun=2e5)
                ))
summary(fearful_dots_m)
```

### Bars

```{r}
fearful_bars_m <- lmer(diff_score~assoc_diff_c+(1+assoc_diff_c|subject_id),
                data=fearful_diff_bars,
                verbose=F,
                control=lmerControl(
                  optimizer="bobyqa",
                  optCtrl=list(maxfun=2e5)
                ))
summary(fearful_bars_m)
```

## Happy

### All

```{r}
happy_m <- lmer(diff_score~assoc_diff_c*plot_c+(1+assoc_diff_c|subject_id),
                data=happy_diff,
                verbose=F,
                control=lmerControl(
                  optimizer="bobyqa",
                  optCtrl=list(maxfun=2e5)
                ))
summary(happy_m)
```

### Dots

```{r}
happy_dots_m <- lmer(diff_score~assoc_diff_c+(1+assoc_diff_c|subject_id),
                data=happy_diff_dots,
                verbose=F,
                control=lmerControl(
                  optimizer="bobyqa",
                  optCtrl=list(maxfun=2e5)
                ))
summary(happy_dots_m)
```

### Bars

```{r}
happy_bars_m <- lmer(diff_score~assoc_diff_c+(1+assoc_diff_c|subject_id),
                data=happy_diff_bars,
                verbose=F,
                control=lmerControl(
                  optimizer="bobyqa",
                  optCtrl=list(maxfun=2e5)
                ))
summary(happy_bars_m)
```

## Sad

### All

```{r}
sad_m <- lmer(diff_score~assoc_diff_c*plot_c+(1+assoc_diff_c|subject_id),
                data=sad_diff,
                verbose=F,
                control=lmerControl(
                  optimizer="bobyqa",
                  optCtrl=list(maxfun=2e5)
                ))
summary(sad_m)
```

### Dots

```{r}
sad_dots_m <- lmer(diff_score~assoc_diff_c+(1+assoc_diff_c|subject_id),
                data=sad_diff_dots,
                verbose=F,
                control=lmerControl(
                  optimizer="bobyqa",
                  optCtrl=list(maxfun=2e5)
                ))
summary(sad_dots_m)
```

### Bars

```{r}
sad_bars_m <- lmer(diff_score~assoc_diff_c+(1+assoc_diff_c|subject_id),
                data=sad_diff_bars,
                verbose=F,
                control=lmerControl(
                  optimizer="bobyqa",
                  optCtrl=list(maxfun=2e5)
                ))
summary(sad_bars_m)
```
